---
title: "R Introduction"
author: "Qahhar"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction: Install Packages
```{r}
#PalmerPenguins
install.packages('palmerpenguins')
```

```{r}
#Load Package

library('palmerpenguins')
```
citation()
citation("palmerpenguins") 

### Python in R Markdown: Setting up environment
```{r}
install.packages("reticulate")

```
```{r}
library(reticulate)

py_env <- virtualenv_create()

# py_install("pip")
# py_install("numpy")
# py_install("pandas")
# py_install("matplotlib")
# py_install("scipy")
# py_install("scikit-learn")
# py_install("seaborn")
# Add more libraries as needed

```
#### Using Python for easy for simple algebra equations
```{python}
#!pip install numpy

```

```{python}
list = [1,2,3,4,5]
import numpy as np
np.mean(list)
```

## Loading data in R from the palmer penguins library
#### Similarly to the load iris from seaborn package in Python
```{r}
data(package = 'palmerpenguins')
library('palmerpenguins')
```

```{r}
data(penguins)
#data(penguins_raw)
```

#View data
```{r}
head(penguins)
#head(penguins_raw)
#View(penguins)
```

#### Describe your data
```{r}
str(penguins)

summary(penguins) #offers a quick descriptive summary of the columns in our dataset
```

#### Play around with the data, Visualization
```{r}
#install.packages("tidyverse")

library(tidyverse)
ggplot(data = penguins, aes(x = flipper_length_mm, y = body_mass_g )) + geom_point(aes(color = species, shape = species ))

```

# what happens when I make new commands

