---
title: 'Base R: HackBio Wk 1'
author: "Qahhar"
date: "2023-10-30"
output: pdf_document
---

The HackBio _DataScience4Life_ contest required it's participant to complete a weekly task

"In this section, you are provided with a dataset and final figures. Your task is to reproduce the figures using the dataset alone". These words mark the exact instructions to be followed for this aspect of the contest

We have also been advised to "Use only base R functions. Do not use any library or package such as ggplot2 to solve the tasks" 

In that light, here is my approach:

# PROBLEM 1: FIG 1 A-E

### Firstly:
Ensure you're working in your preferred directory
```{r}
getwd()
```

# Secondly
We retrieve the dataset to be worked on, in said directory

```{r}
url <- "https://raw.githubusercontent.com/HackBio-Internship/public_datasets/main/R/datasets/Contests/Viz/datasets/fig_One_a_e.dat"

#Downloading the dataset, from the url , and using destfile to assign the file name 
download.file(url, destfile = "fig_1_a_e.dat")
```
### Thirdly:

Read the retrieved data into your environment.
To read in this file, we utilize the read.table function, given a few arguments

 - fig_1_a_e.dat: The name of our dataset
 
 - header: This command when sets to true(T) keeps the header of any given txt file when being read in as a table or a data-frame
 
 - sep: Every plain txt file to be read into R's console has a delimeter, The very thing that defines their structure, For our .dat file, the delimeter is a tab, thus the '\t' notation
 
```{r}
fig_data <- read.table("fig_1_a_e.dat", header= T, sep = '\t')

head(fig_data)
```


### Dataset Previewing
Understanding the dimensions and structure of the dataset being worked with gives the analyst as good idea on what to expect as he/she goes down on the process

```{r}
dim(fig_data) # Returns the dimension of our tibble
```

```{r}
colnames(fig_data) #This returns the no of columns in the given dataset
```
```{r}
summary(fig_data) #Returns a descriptive summary of the entire dataset
```

We get a statistical overview of the dataset. 

- A majority of the  columns are of number types with only one column/variable being a character/string type denoted by **chr**

In order to explore the 'tech' column, we can use "count" to give us a high end description of the columns values
```{r}
length(unique(fig_data$tech))

library(dplyr)
summarize(group_by(fig_data, tech), N = n())
```
From this, we find that the tech column has 10 unique values each of which are 8 in number, in line with the length of the dataset.

**Plotting Figure 1**

As stated previously, the task is to replicate template graphs.
Here is the [link](https://github.com/HackBio-Internship/public_datasets/blob/main/R/datasets/Contests/Viz/figure1a-1f.pdf) for reference

**Setting up a theme e.g color, font size and whatnot saves a ton of stress**
```{r}
tech_colors <- c("#00FFFF","#0066FF","#3300FF","#CC00FF","#FF0099","#FF0000","#FF9900","#CCFF00","#33FF00","#00FF66") #Setting the colors of the unique labels of each group that are subjecting to the plot arguments
```

**We would also be dealing with the tech column alot in this problem, so it's best to store the unique values somewhere
```{r}
technology <- c(unique(fig_data$tech)) # This extracts the unique values of tech in which we are interested in.
```

We would also create a layout for subsequent plots to fit into, and then have an output of a single panel file
```{r}
# Set up a 2x3 layout
par(mfrow = c(2, 3))


#Fig a
#setting up a blank plot
 plot(0,
     type = "n",
     xlim = c(5,40),
     ylim = c(0,12000),
     xlab= "Sequencing Depth",
     ylab = "Unique Fragments in Peaks",
     main = "Fig 1a")

for (i in 1:10) {
  filtered_data <- fig_data[fig_data$tech == technology[i],] 
 # This command basically filters a dataset where tech would be equal to i index in technology
  
  #her telling progam to assign the values accordingly and set the colors individually
  lines(filtered_data$depth,filtered_data$Unique_nr_frag_in_regions, col = tech_colors[i], lwd = 2)
}

# Fig B
#setting up a blank plot
plot(0,
     type = "n",
     xlim = c(5,40),
     ylim = c(0,40),
     xlab= "Sequencing Depth",
     ylab = "Unique Fragments in Peaks",
     main = "Fig 1b")

for (i in 1:10) {
  filtered_data <- fig_data[fig_data$tech == technology[i],] 
 # This command basically filters a dataset where tech would be equal to i index in technology
  
  #her telling progam to assign the values accordingly and set the colors individually
  lines(filtered_data$depth,filtered_data$TSS_enrichment, col = tech_colors[i], lwd = 2)
}

#Fig C
 plot(0,
     type = "n",
     xlim = c(5,40),
     ylim = c(0,0.5),
     xlab= "Sequencing Depth",
     ylab = "Sequencing Effeciency",
     main = "Fig 1c")

for (i in 1:10) {
  filtered_data <- fig_data[fig_data$tech == technology[i],] 
 # This command basically filters the dataset where tech would be equal to i index in technology
  
  #her telling progam to assign the values accordingly and set the colors individually
  lines(filtered_data$depth,filtered_data$X._unique_nr_frag_in_regions_in_cells, col = tech_colors[i], lwd = 2)
}

# Fig D
#setting up a blank plot
 plot(0,
     type = "n",
     xlim = c(5,40),
     ylim = c(0,0.85),
     xlab= "Sequencing Depth",
     ylab = "Median Seurat Score",
     main = "Fig 1d")

for (i in 1:10) {
  filtered_data <- fig_data[fig_data$tech == technology[i],] 
 # This command basically filters the dataset where tech would be equal to i index in technology
  
  #her telling progam to assign the values accordingly and set the colors individually
  lines(filtered_data$depth,filtered_data$median_cell_type_pred_score, col = tech_colors[i], lwd = 2)
}

# Fig E
 plot(0,
     type = "n",
     xlim = c(5,40),
     ylim = c(0,40),
     xlab= "Sequencing Depth",
     ylab = "BCell Strength (FC)",
     main = "Fig 1e")

for (i in 1:10) {
  filtered_data <- fig_data[fig_data$tech == technology[i],] 
 # This command basically filters the dataset where tech would be equal to i index in technology
  
  #her telling progam to assign the values accordingly and set the colors individually
  lines(filtered_data$depth,filtered_data$fc__B_cell, col = tech_colors[i], lwd = 2)
}

```

 Putting all the plots together, it is easier to draw comparisons from the relationships/trend of sequencing depth and other variables in the data frame
 

# PROBLEM 2

These series of problems require datasets to be retrieved

```{r}
#For Figure F
url <- "https://github.com/HackBio-Internship/public_datasets/raw/main/R/datasets/Contests/Viz/datasets/fig_one_f"

# Read in the table
df_f <- read.table(url, header = TRUE, sep = "\t")

# For Figure G
url <- "https://github.com/HackBio-Internship/public_datasets/raw/main/R/datasets/Contests/Viz/datasets/fig_one_g"

df_g <- read.table(url, header = TRUE, sep = "\t")

# For Figure H
url <- "https://raw.githubusercontent.com/HackBio-Internship/public_datasets/main/R/datasets/Contests/Viz/datasets/fig_one_h"

df_h <- read.table(url, header = TRUE, sep = "\t")

# For Figure J
url <- "https://raw.githubusercontent.com/HackBio-Internship/public_datasets/main/R/datasets/Contests/Viz/datasets/fig_one_j"

df_j <- read.table(url, header = TRUE, sep = "\t")

```

We would be using a different color theme, custom
```{r}
new_tech_colors <- c(lightblue = '#5899DA',
               lightorange = '#E8743B',
               lightgreen = "#19A979",
               lightred = "#ED4A7B",
               lightviolet = '#945ECF',
               lightturq = "#13A4B4",
               lightpurple = "#525DF4",
               lightmagenta = "#BF399E",
               grey = "#6C8893",
               lightamber = "#EE6868",
               deepblue = "#2F6497")
```

**Plotting figure 2**
```{r}
# Set up a 2x3 layout
par(mfrow = c(2, 3))


# Fig F
technology_f <- c(unique(df_f$technology)) # This extracts the unique values of tech in which we are interested in.


plot(0,
     type = "n",
     xlim = c(0, 13000),
     ylim = c(0.08, 0.20),
     xlab= "Median Unique Fragments",
     ylab = "Median Scrublet Score",
     main = "Fig 1f",
     las = 1)


for (i in 1:10) {
  filtered_f<- df_f[df_f$technology == technology_f[i],] 
 # This command basically filters the dataset where tech would be equal to i index in technology

  points(filtered_f$Median_Unique_nr_frag_in_regions,
         filtered_f$Mean_scrublet_doublet_scores_fragments, col = new_tech_colors[i], pch = 20)
}

#Fig G

technology_g <- c(unique(df_g$technology)) # This extracts the unique values of tech in which we are interested in.


 plot(0,
     type = "n",
     xlim = c(0, 25000),
     ylim = c(0, 400),
     xlab= "Median Unique Fragments",
     ylab = "Median Freemuxlet",
     main = "Fig 1g",
     las = 1)


for (i in 1:10) {
  filtered_g<- na.omit(df_g[df_g$technology == technology_g[i],]) 
 # This command basically filters the dataset where tech would be equal to i index in technology

  points(filtered_g$Median_unique_nr_frag,
         filtered_g$fmx_delta_donor_llk, col = new_tech_colors[i], pch = 20)
}
 
# Figure H
technology_h <- c(unique(df_h$technology)) # This extracts the unique values of tech in which we are interested in.


 plot(0,
     type = "n",
     xlim = c(2.5, 4.0),
     ylim = c(0.4, 0.9),
     xlab= "Median Unique Fragments",
     ylab = "Median Seurat Score",
     main = "Fig 1h",
     las = 1)


for (i in 1:10) {
  filtered_h<- df_h[df_h$technology == technology_h[i],]
 # This command basically filters the dataset where tech would be equal to i index in technology

  points(filtered_h$log_median_unique_nr_frag_in_regions,
         filtered_h$seurat_score, col = new_tech_colors[i], pch = 20)
}
 
# For Figure J
unq_tech <- unique(df_j$technology)

#blank plot
plot(1,
    type = "n",
    xlim = c(12 , 24.9),
    ylim = c(1.0,4.0),
    xlab = "Median Strength",
    ylab = "Median Distance to TSS",
    main = "Fig 1j",
    las=1)

# grid()

for (i in 1:10) {
 filtered_j <- df_j[df_j$technology == unq_tech[i],]
  
#Checking if the value in 'i'th position in technology matches to the same in the dataframe
 
# Telling program to assign the values accordingly and set the colors individually

points(filtered_j$top2kdars_median_dar_tss_dist/1000,filtered_j$top2kdars_median_dar_logfc,col = new_tech_colors[i], pch=20)

}

``` 


# PROBLEM 3
What this hopes to do is to plot the sample id against variable "Seurat Score", subsetting each plot by the mode of "tech" applicable to the data 

Let's first retrieve the data
```{r}

url <- "https://raw.githubusercontent.com/HackBio-Internship/public_datasets/main/R/datasets/Contests/Viz/datasets/figure_one_k"

df_k <- read.table(url, header = TRUE, sep = "\t")
```
There seems to be an unnecessary data so it would be beneficial to drop it
```{r}
new_data <- df_k[, -1]
head(new_data)
```
Basically what the code above states is that, pick all rows of the previous df_k dataframe and do not select the first column using the (-) and index (1)

_Note: - is not minus_

**Plotting Figure 3**
```{r}
unq_tech <- unique(new_data$tech)
unq_tech


# Accessing the different "Data_for_" objects

for (tech in unq_tech) {
tech_data <- new_data[new_data$tech == tech, ] # Filter data for the current technology
assign(paste("Data_for_", tech, sep = ""), tech_data)
}

for (tech in unq_tech) {
  var_name <- paste("Data_for_", tech, sep = "")
  tech_data <- get(var_name)
  
  print(tech_data)
  
  # Print the first few rows of each object
  cat("Data for", tech, ":\n")
  print(head(tech_data), n = 2)
}

par(mfrow = c(2,5)) #the first value is for rows and the second value is for columns
boxplot(Data_for_ddseq$seurat_cell_type_pred_score ~ Data_for_ddseq$sample_id,
main = "ddseq",
ylim = c(0.0, 1.0),
xlab = "sample_id",
ylab = "Seurat Score",
col = "#1fa342",
xaxt = "n") #xaxt = “n” to remove the ticks from the x-axis

grid()

boxplot (Data_for_mtscatacfacs$seurat_cell_type_pred_score ~
Data_for_mtscatacfacs$sample_id,
main = "mtscatacfacs",
ylim = c(0.0, 1.0),

xlab = "sample_id",
ylab = "Seurat Score",
col = "#a34d1f",
xaxt = "n")

boxplot (Data_for_10xmultiome$seurat_cell_type_pred_score ~
Data_for_10xmultiome$sample_id,
main = "10xmultiome",
ylim = c(0.0, 1.0),
xlab = "sample_id",
ylab = "Seurat Score",
col = "#a3891f",
xaxt = "n")

boxplot (Data_for_10xv11$seurat_cell_type_pred_score ~ Data_for_10xv11$sample_id,
main = "10xv11",
ylim = c(0.0, 1.0),
xlab = "sample_id",
ylab = "Seurat Score",
col = "#ED4A7B",
xaxt = "n")

boxplot (Data_for_10xv11c$seurat_cell_type_pred_score ~ Data_for_10xv11c$sample_id,
main = "10xv11c",
ylim = c(0.0, 1.0),
xlab = "sample_id",
ylab = "Seurat Score",
col = "#42032f",
xaxt = "n")

boxplot (Data_for_10xv2$seurat_cell_type_pred_score ~ Data_for_10xv2$sample_id,
main = "10xv2",

ylim = c(0.0, 1.0),
xlab = "sample_id",
ylab = "Seurat Score",
col = "#210c45",
xaxt = "n")

boxplot (Data_for_hydrop$seurat_cell_type_pred_score ~ Data_for_hydrop$sample_id,
main = "hydrop",
ylim = c(0.0, 1.0),
xlab = "sample_id",
ylab = "Seurat Score",
col = "#0c452e",
xaxt = "n")

boxplot (Data_for_mtscatac$seurat_cell_type_pred_score ~ Data_for_mtscatac$sample_id,
main = "mtscatac",
ylim = c(0.0, 1.0),
xlab = "sample_id",
ylab = "Seurat Score",
col = "#BF399E",
xaxt = "n")

boxplot (Data_for_s3atac$seurat_cell_type_pred_score ~ Data_for_s3atac$sample_id,
main = "s3atac",
ylim = c(0.0, 1.0),
xlab = "sample_id",
ylab = "Seurat Score",
col = "#611918",
xaxt = "n")

boxplot (Data_for_10xv1$seurat_cell_type_pred_score ~ Data_for_10xv1$sample_id,

main = "10xv1",
ylim = c(0.0, 1.0),
xlab = "sample_id",
ylab = "Seurat Score",
col = "#6e6464",
xaxt = "n")
```
































































































